---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# realTimeloads

<!-- badges: start -->
<!-- badges: end -->

The goal of realTimeloads is to compute estimates of analyte flux and load from estimated timeseries of analyte concentration and discharge. An "analyte" is any laboratory measured quantity. Discharge is water volume per unit time. In hydrology, timeseries estimates of analyte concentration are often computed as analyte analysis is cost-prohibitive for continuous water-quality monitoring. A synthetic data set is provided to allow users to explore package functionality and to implement all methods detailed in Livsey (in review)

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
# devtools::install_github("dlivsey/realTimeloads")
``` 

## Example

This is a basic example which shows how to load data from csv files and compute suspended-sediment loads from an acoustic Doppler velocity meter. Data formatted to match the package csv data files can be used to compute loads from user-provided data.

Users are encouraged to explore package functionality via:
vignette('realTimeLoads',package='realTimeLoads') and
?realTimeLoads. Additional worked examples are provided in realTimeloads::ExampleCode() and realTimeloads::ExampleCodeSCI()

A published example using the package methods can be found in: Livsey et al (2020) https://doi.org/10.1007/s12237-020-00734-z


```{r example}
library(realTimeloads)
## basic example code
Input <- realTimeloads::import_data()
Output <- realTimeloads::hADCPLoads(Input)

time <- Output$time
Analyte_flux_timeseries_kt <- Output$Analyte_flux_timeseries_kt
# compute dt (seconds) for used for load integration
dt =  c()
dt[2:length(time)] <- as.numeric(difftime(time[2:length(time)],time[1:length(time)-1],units = "secs"))
dt[1] = median(dt,na.rm=TRUE) # assume time step 1 using median dt

Qspt <- Input$Sediment_Samples$SSCpt_mg_per_liter*
Input$Discharge$Discharge_m_cubed_per_s*dt*1e-9 # actual load (kt) from synthetic data 

plot(time,Analyte_flux_timeseries_kt$median_confidence,
     col='red',type = "l",lwd= 2,xlab = "time (AEST)",ylab="Analyte load (kiloton)",
     main = "Estimated versus actual load")
lines(time,Analyte_flux_timeseries_kt$minus_two_sigma_confidence,
      col='grey')
lines(time,Analyte_flux_timeseries_kt$plus_two_sigma_confidence,
      col='grey')
lines(time,Qspt,col = 'black',lwd= 1.5)
legend("topright",legend = c("Estimated load","Estimation uncertainty","Actual load"),
       lty = c(1,1,1),col = c('red', 'grey', 'black'),pch = c(-1,-1,-1))
       
       
```

